---
title: "HW2"
author: "Chen Mo"
date: "9/28/2020"
output: github_document
---

load library
```{r}
library(tidyverse)
library(readxl)
```

define a path to the dataset
```{r}
path_to_data = "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx"
```

#Problem 1

Read the Mr. Trashwheel dataset

```{r}
 trashwheel_df = 
        read_xlsx(
                path = path_to_data,
                sheet = "Mr. Trash Wheel",
                range = cell_cols("A:N")) %>%
        janitor::clean_names() %>%
        drop_na(dumpster) %>%
        mutate(
                sports_balls = round(sports_balls),
                sports_balls = as.integer(sports_balls)
        )
```

Read the 2017 precipitation data.
```{r}
precip_2017 =
        read_excel(
                path = path_to_data,
                sheet = "2017 Precipitation",
                skip = 1
        ) %>% 
        janitor::clean_names() %>% 
        drop_na(month) %>% 
        mutate(year = 2017) %>%
        relocate(year)
```

Read the 2018 precipitation data.
```{r}
precip_2018 =
        read_excel(
                path = path_to_data,
                sheet = "2018 Precipitation",
                skip = 1
        ) %>% 
        janitor::clean_names() %>% 
        drop_na(month) %>% 
        mutate(year = 2018) %>%
        relocate(year)
```

Combine annual precipitation data.
```{r}
month_df =
        tibble(
                month= 1:12,
                month_name = month.name
        )
precip_df =
        bind_rows(precip_2017, precip_2018)

precip_df =
        left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data. In this dataset:

* The median number of sports balls found in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`
* The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.
